---
title: "Assignment 01"
author: "Mauricio Vazquez & Mariana Luna"
date: '2024-09-27'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Libraries
library(MASS)  
library(ggplot2)
```

### Ex. 8.2: Modern Multivariate Statistical Techniques (Izenman)

Consider the wine data. Compute a LDA, draw a 2D-scatterplot of the first two LDF coordinates, and color-code the points by wine type. What do you notice?

```{r cars}
# Loading dataset
load("C:/Users/mauva/OneDrive/Documents/ITAM/9no Semestre/METODOS MULTIVARIADOS/REPOSITORIO/Multivariate_Statistical_Course_Assignments_Fall2024/ASSIGNMENT_01/wine.rda")

# Parsing to dataframe
wine_df <- as.data.frame(wine)

# Checking head dataset
head(wine_df)

# Parsing dataset column 
wine_df$classdigit <- as.factor(wine_df$classdigit)

# Performing LDA 
lda_model <- lda(classdigit ~ Alcohol + MalicAcid + Ash + AlcAsh + Mg + Phenols + Flav + NonFlavPhenols + Proa + Color + Hue + OD + Proline, data = wine_df)

# Projecting the data onto the LDA axes
lda_values <- predict(lda_model)

# Creating a data frame with the LDA components
lda_df <- data.frame(LD1 = lda_values$x[, 1], LD2 = lda_values$x[, 2], WineType = wine_df$class)

# Plotting
ggplot(lda_df, aes(x = LD1, y = LD2, color = WineType)) +
  geom_point(size = 3) +
  labs(title = "Scatterplot of the First Two LDA Components",
       x = "First Discriminant Component (LD1)",
       y = "Second Discriminant Component (LD2)") +
  theme_minimal() +
  theme(legend.title = element_blank())

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
