---
title: "Assignment 02"
author: "Mauricio Vazquez & Mariana Luna"
date: '2024-11-08'
output:
  pdf_document: default
  html_document: default
---

***Repository link: https://github.com/MauricioVazquezM/Multivariate_Statistical_Course_Assignments_Fall2024***

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = TRUE, fig.width=10)

# Libraries
library(MASS)  
library(ggplot2)
library(GGally)
library(mclust)
library(kableExtra)
library(copula)
library(gridExtra)
```

### (Izenman) Ex. 7.1: Modern Multivariate Statistical 

* Generate a random sample of size \( n = 100 \) from a three-dimensional (\( r = 3 \)) Gaussian distribution, where one of the variables has very high variance (relative to the other two). Carry out PCA on these data using the covariance matrix and the correlation matrix. In each case, find the eigenvalues and eigenvectors, draw the scree plot, compute the PC scores, and plot all pairwise PC scores in a matrix plot. Compare results.

```{r fig.align='center', fig.width=6, fig.height=3, echo=FALSE}
# Paso 1: Configuración y generación de datos
set.seed(42) # Para reproducibilidad
n <- 100
mean <- c(0, 0, 0) # Media de la distribución
cov_matrix <- matrix(c(100, 10, 10, 
                       10, 50, 10, 
                       10, 10, 30), nrow = 3) # Matriz de covarianza

# Generamos los datos aleatorios
data <- mvrnorm(n, mu = mean, Sigma = cov_matrix)
colnames(data) <- c("X1", "X2", "X3")

# Paso 2: PCA con la matriz de covarianza
pca_cov <- prcomp(data, scale. = FALSE) # scale. = FALSE usa la matriz de covarianza
eigenvalues_cov <- pca_cov$sdev^2 # Valores propios
eigenvectors_cov <- pca_cov$rotation # Vectores propios

# Paso 3: PCA con la matriz de correlación
pca_cor <- prcomp(data, scale. = TRUE) # scale. = TRUE usa la matriz de correlación
eigenvalues_cor <- pca_cor$sdev^2 # Valores propios
eigenvectors_cor <- pca_cor$rotation # Vectores propios

# Paso 4: Graficar Scree Plots
par(mfrow=c(1, 2))

# Scree plot para matriz de covarianza
plot(eigenvalues_cov, type="b", xlab="Component", ylab="Eigenvalue",
     main="Scree Plot (Covariance Matrix)")

# Scree plot para matriz de correlación
plot(eigenvalues_cor, type="b", xlab="Component", ylab="Eigenvalue",
     main="Scree Plot (Correlation Matrix)")

# Paso 5: Calcular y graficar los puntajes de las componentes principales
scores_cov <- pca_cov$x # Puntajes de PCA con matriz de covarianza
scores_cor <- pca_cor$x # Puntajes de PCA con matriz de correlación

# Paso 6: Graficar matriz de pares para los puntajes de componentes principales
par(mfrow=c(1, 2))

# Pares de gráficos para matriz de covarianza
pairs(scores_cov, main="Pairwise PC Scores (Covariance Matrix)")

# Pares de gráficos para matriz de correlación
pairs(scores_cor, main="Pairwise PC Scores (Correlation Matrix)")

# Comparación de resultados
cat("Eigenvalues (Covariance Matrix):", eigenvalues_cov, "\n")
cat("Eigenvectors (Covariance Matrix):\n")
print(eigenvectors_cov)

cat("Eigenvalues (Correlation Matrix):", eigenvalues_cor, "\n")
cat("Eigenvectors (Correlation Matrix):\n")
print(eigenvectors_cor)
```

<br>

### (Izenman) Ex. 7.3: Modern Multivariate Statistical Techniques

* In the file `turtles.txt`, there are three variables, length, width, and height, of the carapaces of 48 painted turtles, 24 female and 24 male. Take logarithms of all three variables. Estimate the mean vector and covariance matrix of the male turtles and of the female turtles separately. Find the eigenvalues and eigenvectors of each estimated covariance matrix and carry out a PCA of each data set. Find an expression for the volume of a turtle carapace for males and for females. (Hint: use the fact that the variables are logarithms of the original measurements.) Compare volumes of male and female carapaces.

```{r fig.align='center', fig.width=6, fig.height=3, echo=FALSE}
# Paso 1: Cargar y preparar los datos
load("turtles.rda")
turtles$log_length <- log(turtles$length)
turtles$log_width <- log(turtles$width)
turtles$log_height <- log(turtles$height)

# Dividimos los datos entre machos y hembras
male_turtles <- subset(turtles, sex == "m")
female_turtles <- subset(turtles, sex == "f")

# Paso 2: Calcular el vector de medias y la matriz de covarianza para cada grupo
mean_male <- colMeans(male_turtles[, c("log_length", "log_width", "log_height")])
mean_female <- colMeans(female_turtles[, c("log_length", "log_width", "log_height")])

cov_male <- cov(male_turtles[, c("log_length", "log_width", "log_height")])
cov_female <- cov(female_turtles[, c("log_length", "log_width", "log_height")])

# Paso 3: Realizar PCA en cada conjunto de datos
# PCA para tortugas machos
pca_male <- prcomp(male_turtles[, c("log_length", "log_width", "log_height")], scale. = FALSE)
eigenvalues_male <- pca_male$sdev^2
eigenvectors_male <- pca_male$rotation

# PCA para tortugas hembras
pca_female <- prcomp(female_turtles[, c("log_length", "log_width", "log_height")], scale. = FALSE)
eigenvalues_female <- pca_female$sdev^2
eigenvectors_female <- pca_female$rotation

# Paso 4: Calcular el volumen del caparazón
log_volume_male <- rowSums(male_turtles[, c("log_length", "log_width", "log_height")])
log_volume_female <- rowSums(female_turtles[, c("log_length", "log_width", "log_height")])

mean_volume_male <- exp(mean(log_volume_male))
mean_volume_female <- exp(mean(log_volume_female))

# Paso 5: Imprimir resultados y comparar volúmenes
cat("Vector de medias (machos):", mean_male, "\n")
cat("Vector de medias (hembras):", mean_female, "\n")

cat("Matriz de covarianza (machos):\n")
print(cov_male)
cat("Matriz de covarianza (hembras):\n")
print(cov_female)

cat("Valores propios (machos):", eigenvalues_male, "\n")
cat("Vectores propios (machos):\n")
print(eigenvectors_male)

cat("Valores propios (hembras):", eigenvalues_female, "\n")
cat("Vectores propios (hembras):\n")
print(eigenvectors_female)

cat("Volumen promedio del caparazón (machos):", mean_volume_male, "\n")
cat("Volumen promedio del caparazón (hembras):", mean_volume_female, "\n")

# Comparación de volúmenes
cat("Diferencia de volumen entre machos y hembras:", -(mean_volume_male - mean_volume_female), "\n")
```

<br>

### (Izenman) Ex. 7.10: Modern Multivariate Statistical Techniques

* Consider an \( (r \times r) \) correlation matrix with the same correlation, \( \rho \), say, in the off-diagonal entries. Find the eigenvalues and eigenvectors of this matrix when \( r = 2, 3, 4 \). Generalize your results to any \( r \) variables. As examples, set \( \rho = 0.1, 0.3, 0.5, 0.7, 0.9 \).

```{r fig.align='center', fig.width=6, fig.height=3, echo=FALSE}
# Función para construir la matriz de correlación con correlación constante 'p' en entradas fuera de la diagonal
build_correlation_matrix <- function(r, p) {
  matrix <- matrix(p, nrow = r, ncol = r)
  diag(matrix) <- 1
  return(matrix)
}

p_values <- c(0.1, 0.3, 0.5, 0.7, 0.9)
r_values <- c(2, 3, 4)

# Cálculo de valores y vectores propios para cada combinación de r y p
for (r in r_values) {
  cat("r =", r, "\n")
  for (p in p_values) {
    R <- build_correlation_matrix(r, p)
    
    eigen_result <- eigen(R)
    eigenvalues <- eigen_result$values
    eigenvectors <- eigen_result$vectors
    
    cat("p =", p, "\n")
    cat("Eigenvalues:\n")
    print(eigenvalues)
    cat("Eigenvectors:\n")
    print(eigenvectors)
    cat("\n")
  }
}

# Observación general
cat("Resultado general:\n")
cat("Un valor propio es 1 + (r-1)*p, con el vector propio (1, 1, ..., 1) / sqrt(r).\n")
cat("Los otros (r-1) valores propios son 1 - p, con vectores propios que suman cero.\n")
```

<br>

### (Izenman) Ex. 13.1: Modern Multivariate Statistical Techniques

* Consider the color-stimuli experiment outlined in Section 13.2.1. The similarity ratings are given in the file `color-stimuli` on the book's website. Carry out a classical scaling of the data and show that the solution is a “color circle” ranging from violet (434 mmu) to blue (472 mmu) to green (504 mmu) to yellow (584 mmu) to red (674 mmu). Compare your solution to the nonmetric scaling solution given in Figure 13.3.

```{r fig.align='center', fig.width=6, fig.height=3, echo=FALSE}
# Paso 1: Cargar los datos
load("color.stimuli.rda")

# Paso 2: Convertir similitud a distancia
# Supongamos que la similitud máxima es el valor más alto en los datos
max_similarity <- max(color.stimuli)
distance_matrix <- max_similarity - color.stimuli

# Paso 3: Escalamiento clásico 
mds_result <- cmdscale(distance_matrix, k = 2) # Reducción a 2 dimensiones

# Longitudes de onda aproximadas de los colores en nanómetros (nm)
wavelengths <- c("434 nm (púrpura)", "472 nm (azul)", "504 nm (verde)", 
                 "584 nm (amarillo)", "674 nm (rojo)")

# Paso 4: Graficar el círculo de color
plot(mds_result, type = "n", xlab = "Dimensión 1", ylab = "Dimensión 2", main = "Círculo de Color - Escalamiento Clásico")
text(mds_result, labels = wavelengths, col = c("purple", "blue", "green", "yellow", "red"), cex = 1.2)

# Añadir una circunferencia para resaltar la disposición circular (opcional)
symbols(0, 0, circles = max(sqrt(rowSums(mds_result^2))), add = TRUE, inches = FALSE, lty = 2)
```

<br>

### (Izenman) Ex. 13.2: Modern Multivariate Statistical Techniques

* Consider the Morse-code experiment outlined in Section 13.2.2. The file `Morse-code` on the book's website gives a table of the percentages of times that a signal corresponding to the row label was identified as being the same as the signal corresponding to the column label. A row of this table shows the confusion rate for that particular Morse-code signal when presented *before* each of the column signals, whereas a column of the table shows the confusion rate for that particular signal when presented *after* each of the row signals. This table of confusion rates is not symmetric and the diagonal elements are not each 100%. Now, every square matrix \( M \) can be decomposed uniquely into the sum of two orthogonal matrices, \( M = A + B \), where \( A = \frac{1}{2}(M + M^{T}) \) is symmetric \( (A^{T} = A) \), and \( B = \frac{1}{2}(M - M^{T}) \) is skew-symmetric \( (B^{T} = -B) \) with zero diagonal entries.Find the decomposition for the Mose-code data. Ignore that part of the Morse-code data provided by \( \mathbf{B} \) and carry out a nonmetric scaling only of the symmetric part \( \mathbf{A} \). Decide how many dimensions you think are appropriate for representing the data.

```{r fig.align='center', fig.width=6, fig.height=3, echo=FALSE}
# Paso 1: Cargar los datos
load("morse.rda")

# Paso 2: Descomposición en parte simétrica y antisimétrica
A <- 0.5 * (morse + t(morse))  # Parte simétrica
B <- 0.5 * (morse - t(morse))  # Parte antisimétrica

#cat("Matriz simétrica A:\n")
#print(A)
#cat("\nMatriz antisimétrica B:\n")
#print(B)

# Paso 3: Escalamiento no métrico en A
# Convertir la matriz de similitudes en una matriz de distancias
# (Suponemos que mayores valores indican mayor similitud)
distance_matrix <- max(A) - A
mds_nonmetric <- cmdscale(as.dist(distance_matrix), k = 2, eig = TRUE, add = TRUE)

# Paso 4: Graficar los resultados 
plot(mds_nonmetric$points, type = "n", xlab = "Dimensión 1", ylab = "Dimensión 2", main = "MDS No Métrica para Morse Code (Parte Simétrica)")
text(mds_nonmetric$points, labels = rownames(morse), cex = 1.2)

# Evaluar el número de dimensiones apropiado
# Podemos revisar los eigenvalores para ver cuánta varianza se explica con cada dimensión
cat("Eigenvalores de la MDS no métrica:\n")
print(mds_nonmetric$eig)
```

<br>

### (Izenman) Ex. 13.4: Modern Multivariate Statistical Techniques

* Show that the dissimilarities in the matrix \( \Delta \) are Euclidean distances if and only if the doubly centered matrix \( \mathbf{B} = \mathbf{H} \mathbf{A} \mathbf{H} \) is nonnegative definite, where \( \mathbf{A} \) is given in the classical scaling algorithm of Table 13.5.

```{r fig.align='center', fig.width=6, fig.height=3, echo=FALSE}
# Ejemplo de matriz de disimilitud
A <- matrix(c(0, 1, 2, 1,
              1, 0, 1, 2,
              2, 1, 0, 1,
              1, 2, 1, 0), nrow = 4, byrow = TRUE)

# Construcción de la matriz H
n <- nrow(A)
H <- diag(n) - (1/n) * matrix(1, n, n)

# Matriz doblemente centrada B = HAH
B <- H %*% A %*% H

# Verificar si B es semidefinida positiva
eigenvalues <- eigen(B)$values
cat("Valores propios de B:\n")
print(eigenvalues)
cat("¿B es semidefinida positiva?:", all(eigenvalues >= 0), "\n")

```

<br>

### (Johnson & Wichern) Ex. 8.9: Applied Multivariate Statistical Analysis 

* **Check book**

***Respuesta en pdf anexo***

### (Johnson & Wichern) Ex. 8.12: Applied Multivariate Statistical Analysis 

* Consider the air-pollution data listed in Table 1.5. Your job is to summarize these data in fewer than \( p = 7 \) dimensions if possible. Conduct a principal component analysis of the data using both the covariance matrix \( \mathbf{S} \) and the correlation matrix \( \mathbf{R} \). What have you learned? Does it make any difference which matrix is chosen for analysis? Can the data be summarized in three or fewer dimensions? Can you interpret the principal components?

```{r fig.align='center', fig.width=6, fig.height=3, echo=FALSE}
# Datos del ejercicio
data <- data.frame(
  Total_population = c(5.935, 1.523, 2.599, 4.009, 4.687, 8.044, 2.766, 6.538, 6.451, 3.314, 3.777, 1.530, 2.768, 6.585),
  Median_school_years = c(14.2, 13.1, 12.7, 15.2, 14.7, 15.6, 13.3, 17.0, 12.9, 12.2, 13.0, 13.8, 13.6, 14.9),
  Total_employment = c(2.265, 0.597, 1.237, 1.649, 2.312, 3.641, 1.244, 2.618, 3.147, 1.606, 2.119, 0.798, 1.336, 2.763),
  Health_services_employment = c(2.27, 0.75, 1.11, 0.81, 2.5, 4.51, 1.03, 2.39, 5.52, 2.18, 2.83, 0.84, 1.75, 1.91),
  Median_value_home = c(2.91, 2.62, 1.72, 3.02, 2.22, 2.36, 1.97, 1.85, 2.01, 1.82, 1.80, 4.25, 2.64, 3.17)
)

# Escalar los datos para el PCA usando la matriz de correlación
scaled_data <- scale(data)

# 1. PCA usando la matriz de correlación
pca_corr <- prcomp(scaled_data, scale = TRUE)
summary(pca_corr)

# Varianza explicada por los componentes principales
var_exp_corr <- pca_corr$sdev^2 / sum(pca_corr$sdev^2)
var_exp_corr

# 2. PCA usando la matriz de covarianza
pca_cov <- prcomp(data, scale = FALSE)
summary(pca_cov)

# Varianza explicada por los componentes principales
var_exp_cov <- pca_cov$sdev^2 / sum(pca_cov$sdev^2)
var_exp_cov

# 3. Visualización de los resultados
# Scree Plot
screeplot(pca_corr, type = "lines", main = "Scree Plot (Correlation Matrix)")
screeplot(pca_cov, type = "lines", main = "Scree Plot (Covariance Matrix)")

# Cargas de los componentes principales
loadings_corr <- pca_corr$rotation
loadings_cov <- pca_cov$rotation

print("Cargas usando matriz de correlación:")
print(loadings_corr)

print("Cargas usando matriz de covarianza:")
print(loadings_cov)
```

<br>

### (Johnson & Wichern) Ex. 12.18: Applied Multivariate Statistical Analysis 

* Table 12.12 gives the road distances between 12 Wisconsin cities and cities in neighboring states. Locate the cities in \( q = 1, 2, \) and \( 3 \) dimensions using multidimensional scaling. Plot the minimum stress \( (q) \) versus \( q \) and interpret the graph. Compare the two-dimensional multidimensional scaling configuration with the locations of the cities on a map from an atlas.

```{r fig.align='center', fig.width=6, fig.height=3, echo=FALSE}
# Datos del ejercicio (Tabla 12.12): Matriz de distancias entre ciudades de Wisconsin
distances <- matrix(c(
  0, 107, 127, 141, 129, 230, 232, 210, 340, 180, 334, 325,
  107, 0, 162, 105, 81, 244, 250, 192, 326, 151, 328, 321,
  127, 162, 0, 160, 103, 225, 155, 269, 322, 293, 358, 343,
  141, 105, 160, 0, 57, 266, 251, 181, 300, 95, 278, 261,
  129, 81, 103, 57, 0, 210, 196, 141, 242, 136, 257, 240,
  230, 244, 225, 266, 210, 0, 129, 70, 205, 354, 91, 106,
  232, 250, 155, 251, 196, 129, 0, 199, 105, 324, 53, 68,
  210, 192, 269, 181, 141, 70, 199, 0, 153, 295, 120, 133,
  340, 326, 322, 300, 242, 205, 105, 153, 0, 431, 156, 169,
  180, 151, 293, 95, 136, 354, 324, 295, 431, 0, 389, 368,
  334, 328, 358, 278, 257, 91, 53, 120, 156, 389, 0, 33,
  325, 321, 343, 261, 240, 106, 68, 133, 169, 368, 33, 0
), nrow = 12, ncol = 12, byrow = TRUE)

# Asignamos los nombres de las ciudades
cities <- c("Milwaukee", "Madison", "Green_Bay", "Kenosha", "Racine", "La_Crosse", 
            "Eau_Claire", "Wausau", "Superior", "Beloit", "Rockford", "Chicago")

rownames(distances) <- cities
colnames(distances) <- cities

# 1. Aplicación del MDS para q = 1, 2, 3 dimensiones
mds_1d <- isoMDS(distances, k = 1)
mds_2d <- isoMDS(distances, k = 2)
mds_3d <- isoMDS(distances, k = 3)

# 2. Calcular el stress para q = 1, 2, 3
stress_1d <- mds_1d$stress
stress_2d <- mds_2d$stress
stress_3d <- mds_3d$stress

# Crear un data frame para el gráfico de stress
stress_values <- data.frame(
  q = 1:3,
  stress = c(stress_1d, stress_2d, stress_3d)
)

# Graficar el stress vs. q
ggplot(stress_values, aes(x = q, y = stress)) +
  geom_line() +
  geom_point() +
  labs(title = "Stress vs. Dimensión (q)", x = "Dimensión (q)", y = "Stress") +
  theme_minimal()

# 3. Comparar la configuración bidimensional con el mapa real
mds_2d_points <- data.frame(mds_2d$points)
colnames(mds_2d_points) <- c("Dim1", "Dim2")
mds_2d_points$City <- cities

# Graficar la configuración bidimensional
ggplot(mds_2d_points, aes(x = Dim1, y = Dim2, label = City)) +
  geom_point() +
  geom_text(vjust = -0.5) +
  labs(title = "Configuración MDS Bidimensional de Ciudades",
       x = "Dimensión 1", y = "Dimensión 2") +
  theme_minimal()
```

<br>


### (Johnson & Wichern) Ex. 12.19: Applied Multivariate Statistical Analysis 

* Table 12.13 on page 744 gives the “distances” between certain archaeological sites from different periods, based upon the frequencies of different types of potsherds found at the sites. Given these distances, determine the coordinates of the sites in \( q = 3, 4, \) and \( 5 \) dimensions using multidimensional scaling. Plot the minimum stress \( (q) \) versus \( q \) and interpret the graph. If possible, locate the sites in two dimensions (the first two principal components) using the coordinates for the \( q = 5 \)-dimensional solution. (Treat the sites as variables.) Noting the periods associated with the sites, interpret the two-dimensional configuration.

```{r fig.align='center', fig.width=6, fig.height=3, echo=FALSE}
# Datos del ejercicio (Tabla 12.13): Matriz de "distancias" entre sitios arqueológicos
distances <- matrix(c(
  0, 3.4, 2.9, 2.2, 3.8, 4.6, 4.0, 5.7,
  3.4, 0, 1.2, 3.5, 2.1, 4.5, 3.1, 5.8,
  2.9, 1.2, 0, 2.9, 3.1, 4.2, 3.3, 5.5,
  2.2, 3.5, 2.9, 0, 4.4, 5.1, 4.7, 6.3,
  3.8, 2.1, 3.1, 4.4, 0, 2.8, 1.9, 3.4,
  4.6, 4.5, 4.2, 5.1, 2.8, 0, 2.7, 4.1,
  4.0, 3.1, 3.3, 4.7, 1.9, 2.7, 0, 3.2,
  5.7, 5.8, 5.5, 6.3, 3.4, 4.1, 3.2, 0
), nrow = 8, ncol = 8, byrow = TRUE)

# Asignamos los nombres de los sitios arqueológicos
sites <- c("Site1", "Site2", "Site3", "Site4", "Site5", "Site6", "Site7", "Site8")

rownames(distances) <- sites
colnames(distances) <- sites

# 1. Aplicación del MDS para q = 3, 4, 5 dimensiones
mds_3d <- isoMDS(distances, k = 3)
mds_4d <- isoMDS(distances, k = 4)
mds_5d <- isoMDS(distances, k = 5)

# 2. Calcular el stress para q = 3, 4, 5
stress_3d <- mds_3d$stress
stress_4d <- mds_4d$stress
stress_5d <- mds_5d$stress

# Crear un data frame para el gráfico de stress
stress_values <- data.frame(
  q = 3:5,
  stress = c(stress_3d, stress_4d, stress_5d)
)

# Graficar el stress vs. q
ggplot(stress_values, aes(x = q, y = stress)) +
  geom_line() +
  geom_point() +
  labs(title = "Stress vs. Dimensión (q)", x = "Dimensión (q)", y = "Stress") +
  theme_minimal()

# 3. Proyección a dos dimensiones usando PCA de la solución de q = 5
mds_5d_points <- as.data.frame(mds_5d$points)
colnames(mds_5d_points) <- paste0("Dim", 1:5)

# Aplicamos PCA a los puntos en 5 dimensiones
pca_result <- prcomp(mds_5d_points, scale. = TRUE)
pca_2d <- data.frame(pca_result$x[, 1:2])
pca_2d$Site <- sites

# Graficar la configuración bidimensional
ggplot(pca_2d, aes(x = PC1, y = PC2, label = Site)) +
  geom_point() +
  geom_text(vjust = -0.5) +
  labs(title = "Proyección PCA de la Configuración MDS en 2D",
       x = "Componente Principal 1", y = "Componente Principal 2") +
  theme_minimal()
```

<br>

### (Czado) Ex. 3.3: Analyzing Dependent Data with Vine Copulas

* *Exploratory bivariate copula choices for the seven-dimensional red wine data:* For the data set considered in Exercise 1.7 the pairs plot of the associated pseudo-copula data is given in Fig. 3.15. For each pair of variables propose a pair copula family.

```{r fig.align='center', fig.width=8, fig.height=8, echo=FALSE}
# Leyendo datos
wine_data_path <- "C:/Users/mauva/OneDrive/Documents/ITAM/9no Semestre/METODOS MULTIVARIADOS/REPOSITORIO/Multivariate_Statistical_Course_Assignments_Fall2024/ASSIGNMENT_02/winequality-red.txt"
wine_data <- read.table(wine_data_path, header = TRUE, sep = ";", quote = "\"")

# Cargar y transformar el conjunto de datos wine_data a pseudo-observaciones
wine_data_pseudo <- as.data.frame(sapply(wine_data, function(x) rank(x) / (length(x) + 1)))

# Quedandonos con 7 columnas
wine_data_subset <- wine_data_pseudo[, c("fixed.acidity", "volatile.acidity", "citric.acid", 
                                         "residual.sugar", "chlorides", "free.sulfur.dioxide", 
                                         "total.sulfur.dioxide")]

# Generar un gráfico cópula
ggpairs(
  wine_data_subset,
  lower = list(continuous = wrap("density", alpha = 0.7, color = "blue")),
  diag = list(continuous = wrap("barDiag", color = "grey", bins = 10)),
  upper = list(continuous = wrap("cor", size = 2, color = "red"))
) + 
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    axis.text.y = element_text(angle = 0) 
  )
```

<br>

### (Czado) Ex. 3.4: Analyzing Dependent Data with Vine Copulas

* *URAN3: Exploratory copula choices for the three-dimensional uranium data:* Consider as in Example 2.2 the three-dimensional subset of the *uranium* data set contained in the R package `copula` with variables Cobalt (Co), Titanium (Ti) and Scandium (Sc). As in Example 3.4 transform the original data to the copula scale using marginal empirical distributions. Then explore the empirical normalized contour plots for all pairs of variables and suggest appropriate parametric pair copula families. Check your choices by comparing the fitted to the empirical normalized contour plots.

```{r fig.align='center', fig.width=8, fig.height=8, echo=FALSE}
# Cargando data uranium
data(uranium, package = 'copula')

# Transformar los datos al "copula scale" usando distribuciones empíricas marginales
uranium_pseudo <- as.data.frame(sapply(uranium, function(x) rank(x) / (length(x) + 1)))

# Quedandonos con 3 columnas
uranium_data_subset <- uranium_pseudo[, c("Co", "K","Cs")]

# Grafico copula
ggpairs(
  uranium_data_subset,
  lower = list(continuous = wrap("density", alpha = 0.7, color = "blue")),
  diag = list(continuous = wrap("barDiag", color = "grey", bins = 10)),
  upper = list(continuous = wrap("cor", size = 3, color = "red"))
) +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    axis.text.y = element_text(angle = 0)
  )
```

<br>

### (Czado) Ex. 3.5: Analyzing Dependent Data with Vine Copulas

* *ABALONE3: Exploratory copula choices for the three-dimensional abalone data:* Consider as in Example 2.3 the three-dimensional subset of the *abalone* data set contained in the R package `PivotalR` with variables *shucked*, *viscera*, and *shell*. As in Example 3.4 transform the original data to the copula scale using marginal empirical distributions. Then explore the empirical normalized contour plots for all pairs of variables and suggest appropriate parametric pair copula families. Check your choices by comparing the fitted to the empirical normalized contour plots.

```{r fig.align='center', fig.width=8, fig.height=8, echo=FALSE}
# Path
file_path <- "C:/Users/mauva/OneDrive/Documents/ITAM/9no Semestre/METODOS MULTIVARIADOS/REPOSITORIO/Multivariate_Statistical_Course_Assignments_Fall2024/ASSIGNMENT_02/abalone.data"

# Leyendo el archivo
abalone_data <- read.table(file_path, sep = ",", header = FALSE)

# Transformar los datos al "copula scale" usando distribuciones empíricas marginales
abalone_pseudo <- as.data.frame(sapply(abalone_data, function(x) rank(x) / (length(x) + 1)))

# Grafico copula
ggpairs(
  abalone_pseudo,
  lower = list(continuous = wrap("density", alpha = 0.7, color = "blue")),
  diag = list(continuous = wrap("barDiag", color = "grey", bins = 10)),
  upper = list(continuous = wrap("cor", size = 3, color = "red"))
) +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    axis.text.y = element_text(angle = 0)
  )
```

<br>

### (Czado) Ex. 3.6: Analyzing Dependent Data with Vine Copulas

* *The effect of the degree of freedom in a bivariate Student’s t copula on the contour shapes:* For \( df = 2, \dots, 30 \) draw the normalized contour plots, when the association parameter is \( \rho = .7 \). Do the same for \( \rho = -.2 \). How do these plots change when you fix \( \tau = .7 \) and \( \tau = -.2 \), respectively.

```{r fig.align='center', fig.width=6, fig.height=3, echo=FALSE}

```

<br>

### (Czado) Ex. 3.10: Analyzing Dependent Data with Vine Copulas

* *Conditional distribution of the Clayton copula:* Derive and visualize the \( h \) functions \( C_{2|1}(u_2 | u_1 = .5) \) and \( C_{1|2}(u_1 | u_2 = .5) \) of a bivariate Clayton copula with a Kendall’s \( \tau = .5 \) and \( \tau = .8 \), respectively. Compare the two functions. Do the same for a 90° rotated Clayton copula.

```{r fig.align='center', fig.width=6, fig.height=3, echo=FALSE}

```

<br>